---
title: Homework 1
author: David Lewis
date: January 30, 2025
---


## load the data
```{r}
studentdata <- LearnBayes::studentdata
```

## 1. Size of the data
```{r}
cat(paste("number of rows=", nrow(studentdata), "\nnumber of columns=", ncol(studentdata), sep=""))
```

## 2. Top 6 rows of the data
```{r}
head(studentdata, n=6)
```

## 3.  Pull out the complete data of Students 4, 22, 517, and 533.
```{r}
studentdata[studentdata["Student"] %in% c(4, 22, 517, 533)]
```

## 4. Using the documentation command, describe each variable in the data.
```{r help, message=FALSE}
library(printr)
?LearnBayes::studentdata
```
The description for each variable is under the "format" section of the above output.


## 5. What is the nature of each variable?
```{r}
data.frame("class" = sapply(studentdata, class))
```

## 6. Show the summary statistics of each variable.
```{r}
summary(studentdata)
```

## 7. What is the gender distribution?
```{r}
summary(studentdata$Gender)
```

## 8. What is the most favorite number of the students?
```{r}
sort(table(studentdata$Number), decreasing=TRUE)
```
The most favorite number of the students is 7

## 9. What is the second most favorite number of the students?
The second favorite number is 5.

## 10. What is the least favorite number of students?
The least favorite number is 10.

## 11. Use the ‘table’ command on ‘studentdata$Shoes’ and show the output.
```{r}
table(studentdata$Shoes)
```


## 12. What is unusual about the output in 11?

One interesting thing is that some students seem to have an extra shoe that does not belong to a pair.

These data also appear to be spread out in a strange way. While it is true that the second most common number
of pairs of shoes is 10, and most other students have a similar number of shoes, there are a relatively large number of students
with 20 pairs of shoes.

```{r}
plot(table(studentdata$Shoes), xlab="number of shoes", ylab="number of students")
```

Plotting the data shows that while much of the data fits a gaussian distribution, there are many "spikes" that don't fit the distribution.

## 13. Use the ‘table’ command on ‘studentdata$Drink.’

```{r}
table(studentdata$Drink)

```

## 14. Cross-tabulate ‘Gender’ and ‘Drink.’
```{r}
genderDrink <- table(studentdata$Gender, studentdata$Drink)
```

## 15. Calculate proportions row-wise and column-wise as well and correct to two decimal places.

### Row-wise
```{r}
round(prop.table(genderDrink, 1), 2)
```

### Column-wise
```{r}
round(prop.table(genderDrink, 2), 2)
```