---
title: Exam 1 (part 2)
author: David Lewis
date: March 26, 2025
---

```{r, include=FALSE}
library(printr)
library(tidyverse)
```

# Question 6: Write a function se() which calculates the standard error of its argument x

```{r}
test_cases <- list(c(1, 2, 3, 4, 5), c(3, 5, "a", 7), c(3, NA, 8, 2))

se <- function(x) {
    sd(x) / sqrt(length(x))
}

lapply(test_cases, se)
```

As we see from the test cases, if a non-numeric value (or NA) is present in the vector, then the result of the function is NA.


# Question 7: improve the definition of se() from Question 6

```{r}
se <- function(x, na.rm = FALSE) {
    # Check if non-numeric
    if (!is.numeric(x)) {
        warning("Argument is not numeric: Returning NA")
        return(NA)
    }
    # remove NAs
    if (na.rm) {
        x <- na.omit(x)
    }
    sd(x) / sqrt(length(x))
}

se(c(3, 5, "a", 7))
se(c(3, NA, 8, 2))
se(c(3, NA, 8, 2), na.rm = TRUE)
```

# Question 8

> The built-in data “trees” provide the measurements of diameter (Girth), height and volume of
> trees. These are in inches for Girth, feet for Height and cubic feet for Volume. We will focus on
> Girth and Height. Write a function convert() that converts the values into the metric system
> (meters). Your function should take as arguments the value to convert and the unit it was
> expressed in (inches or feet) (use the following: x inches = x*0.0254 meters, x feet = x*0.3048
> meters). If the first argument is not numeric or if the second argument is not “inch” or “feet”,
> use stop() function with an error message. 

```{r}
convert <- function(x, units) {
    if (!(units == "feet" || units == "inches")) {
        stop('The unit was not "inches" or "feet"')
    }
    if (!is.numeric(x)) {
        stop("The data is not numeric")
    }
    if (units == "inches") {
        return(x * 0.0254)
    }
    # feet
    return(x * 0.3048)
}

convert(8.3, "inches")
convert(8.3, "feet")
```
```{r, error=TRUE}
convert(8.3, "foot")
```

> Use your function to convert the values of Girth and Height to meters. Using cor.test() function,
> calculate the correlation (and the corresponding p-value) between Girth and Height in meters.

```{r}
Height_meters <- convert(trees$Height, "feet")
Girth_meters <- convert(trees$Girth, "inches")

cor.test(Girth_meters, Height_meters)
```

## Question 9 

> Using the built-in chickwts data (contains measured growth rates of chickens in dependence of
> various feed supplements), produce the following figure using ggplot():
> Use the colors 'grey50', 'grey60', 'grey70', 'grey80', 'grey90' and 'grey100' for the bars. Note the
> custom titles and axis labels.

```{r}
ggplot(chickwts, aes(x = feed, y = weight)) +
    geom_boxplot(fill = c("grey50", "grey60", "grey70", "grey80", "grey90", "grey100")) +
    xlab("Chicken weights depending on different feed supplement") +
    ylab("Weight after six weeks (grams)") +
    ggtitle("chickwts data") +
    theme(plot.title = element_text(hjust = 0.5))
```




## Question 10. 
> Using the built-in cars data (contains distances taken to stop from a certain speed). Plot dist as a
> function of speed without titles and without axes (axes=FALSE). Add a line with slope 4 and
> intercept -17.5 (you can use abline()). Furthermore add axes with the command axis().
> Specify the positions of the ticks with at = 'pretty' positions, generated with pretty(speed) and
> pretty(dist) (e.g. at = pretty(cars$speed)), respectively. Add a legend with
> legend(). The text in the legend is generated with the command expression().
> Furthermore add a main title and a subtitle with the command title(main =, sub = ).
> (The command locator(1) might help you to find a good position for the legend).

```{r}
plot(cars$speed, cars$dist, axes = FALSE, col = "red", xlab = "speed", ylab = "distance")
abline(-17.5, 4)
axis(1, at = pretty(cars$speed))
axis(2, at = pretty(cars$dist))
legend(
    x = 5, y = 120,
    expression(paste(beta[0], " = 17", ", ", beta[1], " = 14")), lty = 1
)
title(main = "Distance taken to stop a car at a certain speed", sub = "Distance as a function of speed")
```

