---
title: Exam 1 (Part 1)
author: David Lewis
date: March 25, 2025
---

```{r include=FALSE}
library(printr)
```

> Note that I used special rmarkdown syntax to embed some results directly into my prose.
> you can see the rmarkdown file for more details

# Question 1

```{r}
data <- 90 * 1:100 - (1:100)^2 + 1000
```

## What is the first, the seventeenth and the last entry of the vector data?

The first member of the vector is `r data[1]`.
The 17th member of the vector is `r data[17]`.
The last member of the vector is `r data[length(data)]`


## What is the maximum of the vector data? At which index is the maximum attained?
The maximum member of the vector is `r max(data)`.
The index of this maximum is `r which.max(data)`.

## Plot the vector data with plot(data) and visually confirm your last result.

```{r}
plot(data)
y <- max(data)
x <- which.max(data)

arrows(x, y - 1000, x, y - 10)
```

The arrow points the max entry, confirming the last result.


## At which indices are the entries of data between 2000 and 2500?

Assuming "between" is not inclusive.

```{r}
index <- which(data > 2000 & data < 2500)
```

The indices that are between 2000 and 2500 are `r index` .

# Question 2

```{r}
m <- matrix(11:35, nrow = 5, byrow = TRUE)
```

## What is the entry in the third row and forth column? 

The entry in the third row and 4th column is `r m[3,4]`


## Briefly describe in words what `m[2:4,3:5]` returns.

The syntax `2:4` represents a range of numbers from 2 to 4, i.e (2,3,4). So in this case `m[2:4, 3:5]` is 
a submatrix of m containing the 2,3,4 rows and 3,4,5 columns.

```{r}
m[2:4, 3:5]
```

# Question 3
## Define the variable treatment as a vector of length 100 with elements: (“yes”, “control”, “yes”, “control”, ... , “yes”, “control”)

```{r}
treatment <- rep(c("yes", "control"), 50)
```

## Define the variable smoker as a vector of length 100 with elements: (“yes”, ”yes”, ”no”, ”no”, ”no”, ... , “yes”, “yes”, “no”, “no”, “no”)

```{r}
smoker <- rep(c("yes", "yes", "no", "no", "no"), 20)
```
## Define the vector :
```{r}
lifespan <- abs(round(100 * sin(1:100)))
```

## Create a data frame with treatment, smoker, and lifespan vectors. You may think of lifespan as the life span of the individuals.
```{r}
data <- data.frame("treatment" = treatment, "smoker" = smoker, "lifespan" = lifespan)
```

## Define a new vector x which consists of all elements of lifespan at whose index in smoker is the element "yes". 
```{r}
x <- data[data$smoker == "yes", ]$lifespan
```

## What is the maximum of lifespan over all smokers?
The maximum lifespan over all smokers is `r max(x)`


## Half of the individuals got a certain treatment. Produce a new vector consisting of the lifespans of all individuals which are smokers and got the treatment.

```{r}
lifespan_smokers_with_treatment <- data[data$smoker == "yes" & data$treatment == "yes", ]$lifespan
```


# Question 4
## using the Wages data from Ecdat library (also attached as Wages.csv), do the following: Show the top and bottom six rows of the data.
```{r}
wages <- read.csv("Wages.csv")
wages$X <- NULL
```

```{r}
rbind(head(n = 6, wages), tail(n = 6, wages))
```

## How would you calculate the relative proportion (in percentages) of male and female workers?
```{r}
prop.table(table(wages$sex)) * 100
```


## How do you calculate these proportions separately for workers in the south (yes, no)? 

```{r}
prop.table(table(subset(wages, select = c("south", "sex"))), margin = 1) * 100
```

## How do you calculate these proportions for all workers with an lwage larger than 6.5?

```{r}
x <- wages[wages$lwage > 6.5, ]

prop.table(table(subset(x, select = c("south", "sex"))), margin = 1) * 100
```



# Question 5
> You again use the dataset Wages. Use the command aggregate() to construct a data frame
> with three columns: In one column you have ed, in the second column you have sex, and in the
> third column you have for each possible combination of ed and sex the median of lwage for
> this combination. Since there are only 14 levels of ed and only two levels of sex, your new data
> frame will have at most 28 rows. How many rows does your data frame actually have? Why does
> it have fewer than 28 rows?

```{r}
aggregate_lwage <- aggregate(lwage ~ ed + sex, data = wages, FUN = median)
```


The data frame only has `r nrow(aggregate_lwage)` rows. This is because some combinations of the data do not
exist. In this case (ed:4, sex:female), (ed:5, sex:female) and (ed:6, sex:female) do not exist. This is the reason why there are 3 less rows.
