---
title: Homework 2
author: David Lewis
date: February 20, 2025
---

## Load the data
```{r}
data <- USArrests
```

## 1. What is the size of the data? 2 points
```{r}
dim(data)
```

There are 50 rows and 4 columns

## 2. Describe the data. 2 points
```{r}
library(printr)
?USArrests
```


## 3. Show the top nine rows of the data. 2 points
```{r}
head(data, 9)
```


## 4. Write a function that calculates and returns length, mean and standard deviation of a vector.                                                 2 points          
```{r}
stats <- function(vector) {
    if (!is.vector(vector) || !is.atomic(vector)) {
        stop("Input is not a vector")
    }
    if (sum(is.na(vector)) > 0) {
        message("WARNING: input vector contains NA values, removing")
        vector <- na.omit(vector)
    }

    stat <- list()

    stat$length <- length(vector)
    stat$sd <- sd(vector)
    stat$mean <- mean(vector)

    return(stat)
}
```


## 5. Does you function deal with missing values? 2 points
Yes, the function prints a warning if the input vector contains missing values and removes them before calculating the mean
standard deviation and length.

## 6. How can you modify it so that missing values are ignored? 2 points
There is a check in the function before the calculations that detects if the input vector contains any NA values via `is.na`


## 7. Apply the function from 6 to each column of the USArrests data. Report the means and standard deviations.                                     2 points
```{r}
print_only_mean_and_sd <- function(l) {
    list(mean = l$mean, standard_deviation = l$sd)
}
statistics <- apply(data, 2, stats)
sapply(statistics, print_only_mean_and_sd)
```

## 8. Generate a random sample from a normal distribution with characteristics similar to the murder column in USArrests data (same size, same mean, same sd). 2 points
```{r}
set.seed(0) # so my commentatary on the histograms make sense
sample_like_murder <- rnorm(statistics$Murder$length, statistics$Murder$mean, statistics$Murder$sd)
```

## 9. Create two histograms - for the simulated and the original murder data. Use par(mfrow = c(2,1)) command to plot two histograms in one window. Do the histograms look similar?                                                             2 points

```{r}
par(mfrow = c(2, 1))
hist(sample_like_murder, main = "Simulated Murder")
hist(data$Murder, main = "Murder")
```

They do not look that similar, the `rnorm` sample resembles a normal distribution, while the actual murder histogram is much more irregular when compared to a normal distribution.
However, there are not enough data points (50) to really make a valid assumption about the "normal-ness" of the data.

## 10. Create a new column called MurderRatio – the ratio of murder to urban population percent. Use with() or within() functions. 2 points
```{r}
data <- within(data, {
    MurderRatio <- Murder / UrbanPop
})
```

## 11. Define a new variable called MurderGroup as follows - if MurderRatio is higher than 0.25, the state is considered "High", if MurderRatio is between 0.05 and 0.25, the state is considered "Medium", and the state is "Low" otherwise. 2 points
```{r}
MurderGroup <- cut(data$MurderRatio, breaks = c(-Inf, 0.05, 0.25, Inf), labels = c("Low", "Medium", "High"))
```

## 12. What percent of the states is in High, Low, Medium groups? 2 points
```{r}
percent <- function(vector) {
    (table(vector) / length(vector)) * 100
}
percent(MurderGroup)
```

## 13. Generate a random sample from the set of labels c(“High”, “Low”, “Medium”) with probabilities from Question 12. Use the documentation command to learn how to specify the probabilities of selection. What percent of the states are in High, Low, and Medium groups in simulated data? Why are these percentages not equal to percentages in 11? 4 points
```{r}
s <- sample(c("High", "Low", "Medium"), size = 50, replace = TRUE, prob = c(0.2, 0.72, 0.08))
percent(s)
```

Because the sample size is small, the sample is likely to not exactly represent the distribution (given by the probabilities.) Increasing the sample size should better fit the probabilities.

```{r}
set.seed(0)
s <- sample(c("High", "Low", "Medium"), size = 5000, replace = TRUE, prob = c(0.2, 0.72, 0.08))
percent(s)
```
When the sample size is increased to 5000, the percentages are a much closer to the given probabilities.

## 14. Summarize all variables in the USArrests data by MurderGroup by group means? Do you see a pattern in different crime counts by groups? 2 points

```{r}
aggregate(. ~ MurderGroup, data = data, FUN = mean)
```

The highest rate of murders occur in populations with a low ratio of urban population. There is a similar pattern that occurs with the other crimes in relation to UrbanPopulation There is a similar pattern that occurs with the other crimes in relation to UrbanPopulation. 
The only contrary variable is Rape, which occurs most often at the Medium murder ratio (the highest urban population).

## 15. Optional problem for extra points: 5 points
   Results of 100 coin toss experiments are recorded as follows: 0 = “Tail”, 1 = “Head”. The output, X, is a string of 0’s and 1 of length 100. Then, the number of times when we see 1-0-0 in X is calculated and equal to 21.  (Example: if X = c(0001001110100), 1-0-0 occurs 2 times).  Do you believe that this is a fair coin (probability of 1 = 0.5)? Justify your answer.
   HINTS: You may use other functions, or functions below:
   
   - You can use paste(x, collapse = "") to convert x to a string;
   - You can use unlist(strsplit(x, split = "")) to convert string x to a vector of characters;
   - gsub("100", "X", x) can be used to replace a “100” pattern with “X” pattern.

```{r}
matches100 <- function(seed = 0) {
    set.seed(seed)
    s <- sample(c("0", "1"), size = 100, replace = TRUE, prob = c(0.5, 0.5))
    s <- paste(s, collapse = "")
    l <- gregexpr("100", as.vector(s))[[1]] ## we can use gregexpr to count all "100" matches
    length(l)
}
samples <- sapply(1:100000, matches100)
par(mfrow = c(1, 1))
hist(samples)
```


As you can see from the histogram, 21 "100" matches is extremely unlikely given 100000 trials of 100 coin tosses.

To see just how unlikely, we can count the max number of "100" matches.K


```{r}
max(samples)

length(which(samples == max(samples)))
```

As we can see, the number of "100" matches is only 21 times out of 100000 trials. 


Verdict: While it is possible that it is a fair coin, it is very unlikely.